Estimation of Stochastic Wilson and Cowan Neural Field Model using Expectation Maximisation Algorithm

### 2011-08-26 ltx/data_paper/DataPaper.tex
edits to discussion

While the EKS has been applied in the EM algorithm for learning nonlinear models \cite{Ghahramani1999}, this is the first work (to the best author's knowledge) where the URTSS is used in this context. However, as in the case of the EKS, the stability of the algorithm cannot be ensured. This is because the EKS or URTSS provides an approximation to the exact E-step. Following this, the maximisation of the approximated lower bound is not guaranteed to increase the log-likelihood function. Nevertheless, it is envisaged that the use of the URTSS will provide a more stable algorithm due to its superior performance.

Typically a sequence of likelihood values converge to some stationary points i.e. global (local) maximum or a saddle point. If the sequence of the EM is trapped in a saddle point a small random perturbation diverges the algorithm from such a stationary value \cite{McLachlan1997}. In general convergence of the EM algorithm to any stationary points either global (local) maxima or saddle points depends on the initialisation. Here a bounded sequence of the state vectors were used to initialise the algorithm which gives a satisfactory convergence and good state and parameter estimations.

Since the non-linearity is a sigmoid function first-order Taylor linearisation provides a good approximation to the lower bound of the likelihood function, however in general settings with more complicated non-linear behaviour higher order terms might need to be included, this however introduces intractable computational complexity.

>>>>

While the EKS has been applied in the EM algorithm for learning nonlinear models \cite{Ghahramani1999}, this is the first work (to the best author's knowledge) where the URTSS is used in this context. However, as in the case of the EKS, the stability of the algorithm cannot be ensured. This is because the EKS and the URTSS provide an approximation to the exact E-step. Following this, the maximisation of the approximated lower bound is not guaranteed to increase the log-likelihood function. Nevertheless, it is envisaged that the use of the EM algorithm using the URTSS will provide a more stable algorithm due to the superior performance of the UKF over the EKF.

Typically, a sequence of likelihood values will converge to a stationary point i.e. global (local) maximum or a saddle point. If the sequence of the EM is trapped in a saddle point, a small random perturbation diverges the algorithm from the local maxima \cite{McLachlan1997}. In general, convergence of the EM algorithm to any stationary point (ie. global maxima, local maxima or saddle point) depends on the initialisation. In the examples presented in this paper, a bounded sequence of the state vectors was used to initialise the algorithm, which gave a satisfactory convergence and good state and parameter estimations.

The first-order Taylor expansion provides a good local approximation of the sigmoid function, thus the EM provides a good approximation to the lower bound of the likelihood function. However, in a general settings with a more complicated nonlinearity, higher-order terms may be required to provide a good approximation. However, using higher-order terms introduces intractable computational complexity.
